---
title: "Simulation with two breeding goals"
author: "Martin Johnsson"
output: html_document
---


## Housekeeping

Load packages and simulation results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r}
library(AlphaSimR)
library(dplyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(tidyr)

source("R/simulation_functions.R")
```

```{r}
results_avoid_cxc_snp <-
    read_results(system("ls simulations/divergence_simulations/avoid_carrier_x_carrier_snp/results_*.Rds",
                        intern = TRUE))
results_avoid_cxc_qtl <- 
    read_results(system("ls simulations/divergence_simulations/avoid_carrier_x_carrier_qtl/results_*.Rds",
                        intern = TRUE))

results_unknown_snp <- 
    read_results(system("ls simulations/divergence_simulations/unknown_lethal_snp/results_*.Rds",
                        intern = TRUE))
results_unknown_qtl <- 
    read_results(system("ls simulations/divergence_simulations/unknown_lethal_qtl/results_*.Rds",
                        intern = TRUE))

results_selection_snp <- 
    read_results(system("ls simulations/divergence_simulations/selection_against_snp_0exempt/results_*.Rds",
                        intern = TRUE))
results_selection_qtl <- 
    read_results(system("ls simulations/divergence_simulations/selection_against_qtl_0exempt/results_*.Rds",
                        intern = TRUE))

results_selection_snp_10exempt <-
    read_results(system("ls simulations/divergence_simulations/selection_against_snp_10exempt/results_*.Rds",
                        intern = TRUE))
results_selection_qtl_10exempt <-
    read_results(system("ls simulations/divergence_simulations/selection_against_qtl_10exempt/results_*.Rds",
                        intern = TRUE))

results_selection_snp_100exempt <- 
    read_results(system("ls simulations/divergence_simulations/selection_against_snp_100exempt/results_*.Rds",
                        intern = TRUE))
results_selection_qtl_100exempt <- 
    read_results(system("ls simulations/divergence_simulations/selection_against_qtl_100exempt/results_*.Rds",
                        intern = TRUE))
```

Generation counter

```{r}
generations <- 1:20
```



## Basic features of the simulation

### Genetic mean trend

```{r}
stats_avoid_cxc_snp <- combined_stats(results_avoid_cxc_snp)
stats_avoid_cxc_qtl <- combined_stats(results_avoid_cxc_qtl)

stats_unknown_snp <- combined_stats(results_unknown_snp)
stats_unknown_qtl <- combined_stats(results_unknown_qtl)

stats_selection_snp <- combined_stats(results_selection_snp)
stats_selection_qtl <- combined_stats(results_selection_qtl)

stats_selection_snp_10exempt <- combined_stats(results_selection_snp_10exempt)
stats_selection_qtl_10exempt <- combined_stats(results_selection_qtl_10exempt)

stats_selection_snp_100exempt <- combined_stats(results_selection_snp_100exempt)
stats_selection_qtl_100exempt <- combined_stats(results_selection_qtl_100exempt)
```



### Evolution of genetic correlation

```{r}
average_genetic_correlation <- summarise(group_by(stats_avoid_cxc_snp, generation),
                                         average = mean(cor_g),
                                         upper = quantile(cor_g, 0.95),
                                         lower = quantile(cor_g, 0.05))

plot_genetic_correlation <- ggplot() +
  geom_line(aes(x = generation,
                y = cor_g,
                group = replicate),
            colour = "grey",
            data = stats_avoid_cxc_snp) +
  geom_pointrange(aes(x = generation,
                      ymin = lower,
                      y = average,
                      ymax = upper),
                  data = average_genetic_correlation) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Generation") +
  ylab("Genetic correlation between breeding goals")

plot_genetic_correlation
```


```{r}
pdf("figures/supplementary_fig_genetic_correlation.pdf")
print(plot_genetic_correlation)
dev.off()
```



## Allele frequency changes

### Lethal allele frequency

```{r}
carriers_avoid_cxc_snp <- combined_carriers(results_avoid_cxc_snp)
carriers_avoid_cxc_qtl <- combined_carriers(results_avoid_cxc_qtl)

carriers_unknown_snp <- combined_carriers(results_unknown_snp)
carriers_unknown_qtl <- combined_carriers(results_unknown_qtl)

carriers_selection_snp <- combined_carriers(results_selection_snp)
carriers_selection_qtl <- combined_carriers(results_selection_qtl)

freq_formatting <- list(theme(legend.position = "none"),
                        ylim(0, 0.2))

plot_freq <- ((lethal_frequency_plot(carriers_avoid_cxc_snp) +
                 ggtitle("Lethal allele frequency change without balancing selection") +
                 freq_formatting) |
                (lethal_frequency_plot(carriers_avoid_cxc_qtl) +
                   ggtitle("Lethal allele frequency change with balancing selection") +
                   freq_formatting)) /
  ((lethal_frequency_plot(carriers_unknown_snp) +
      ggtitle("Unknown lethal, without balancing selection") +
      freq_formatting) |
     (lethal_frequency_plot(carriers_unknown_qtl) +
        ggtitle("Unknown lethal, with balancing selection") +
        freq_formatting)) /
  ((lethal_frequency_plot(carriers_selection_snp) +
      ggtitle("Culling in sires, without balancing selection") +
      freq_formatting) |
     (lethal_frequency_plot(carriers_selection_qtl) +
        ggtitle("Culling in sires, with balancing selection") +
        freq_formatting))

plot_freq    
```




### Frequency at the end of the simulation

```{r}
end_carriers_avoid_cxc_snp <- get_end_carriers(carriers_avoid_cxc_snp)
end_carriers_avoid_cxc_qtl <- get_end_carriers(carriers_avoid_cxc_qtl)

end_carriers_unknown_snp <- get_end_carriers(carriers_unknown_snp)
end_carriers_unknown_qtl <- get_end_carriers(carriers_unknown_qtl)

end_carriers_selection_snp <- get_end_carriers(carriers_selection_snp)
end_carriers_selection_qtl <- get_end_carriers(carriers_selection_qtl)

carrier_stats <- rbind(transform(end_carrier_stats(end_carriers_avoid_cxc_snp),
                                 case = "Baseline"),
                       transform(end_carrier_stats(end_carriers_avoid_cxc_qtl),
                                 case = "Baseline, balancing"),
                       
                       transform(end_carrier_stats(end_carriers_unknown_snp),
                                 case = "Unknown lethal"),
                       transform(end_carrier_stats(end_carriers_unknown_qtl),
                                 case = "Unknown lethal, balancing"),
                       
                       transform(end_carrier_stats(end_carriers_selection_snp),
                                 case = "Culling sires"),
                       transform(end_carrier_stats(end_carriers_selection_qtl),
                                 case = "Culling sires, balancing"))

plot_end_freq <- qplot(x = case, ymin = lower, ymax = upper, y = mean_frequency,
                       data = carrier_stats,
                       geom = "pointrange") +
  ggtitle("Lethal frequency in generation 16-20 (5% quantile, mean, 95% quantile)")

plot_end_freq
```

Mean of carrier frequency in generation 20:

```{r}
mean(filter(carriers_avoid_cxc_snp, generation == 20)$carriers/6000)
mean(filter(carriers_avoid_cxc_qtl, generation == 20)$carriers/6000)

mean(filter(carriers_unknown_snp, generation == 20)$carriers/6000)
mean(filter(carriers_unknown_qtl, generation == 20)$carriers/6000)
```


Standard deviation of carrier frequency in generation 20:

```{r}
sd(filter(carriers_avoid_cxc_snp, generation == 20)$carriers/6000)
sd(filter(carriers_avoid_cxc_qtl, generation == 20)$carriers/6000)

sd(filter(carriers_unknown_snp, generation == 20)$carriers/6000)
sd(filter(carriers_unknown_qtl, generation == 20)$carriers/6000)
```


Frequency in selection scenarios in generation 5:

```{r}
mean(filter(carriers_selection_snp, generation == 5)$carriers)/6000/2
mean(filter(carriers_selection_qtl, generation == 5)$carriers)/6000/2
```



## Simulation figure 1: Lethal allele frequency for (un)known lethal

with and without balancing selection

```{r}
carriers_combined <- rbind(transform(carriers_avoid_cxc_snp,
                                     case = "avoid carrier-carrier matings",
                                     selection = "no balancing selection"),
                           transform(carriers_avoid_cxc_qtl,
                                     case = "avoid carrier-carrier matings",
                                     selection = "balancing selection"),
                           transform(carriers_unknown_snp,
                                     case = "no testing",
                                     selection = "no balancing selection"),
                           transform(carriers_unknown_qtl,
                                     case = "no testing",
                                     selection = "balancing selection"))

carriers_combined_summary <- summarise(group_by(carriers_combined, generation, case, selection),
                                       average = mean(carriers/n),
                                       lower = quantile(carriers/n, 0.05),
                                       upper = quantile(carriers/n, 0.95))

fig1_carrier_frequency <- ggplot() +
    geom_line(aes(x = generation,
                  y = carriers/n,
                  group = paste(replicate, case, selection)),
              colour = "grey",
              data = carriers_combined) +
    geom_pointrange(aes(x = generation,
                        y = average,
                        ymin = lower,
                        ymax = upper),
                    data = carriers_combined_summary) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          strip.background = element_blank()) +
    facet_grid(case ~ selection) +
    ggtitle("Carrier frequency with and without balancing selection") +
    ylab("Carrier frequency") +
    xlab("Generations")

fig1_carrier_frequency
```

```{r}
pdf("figures/fig1_carrier_frequency.pdf")
print(fig1_carrier_frequency)
dev.off()
```

Summary of carriers in generation 20:

```{r}
filter(carriers_combined_summary, generation == 20)
```



## Number of affected individuals (unknown lethal case)


```{r}
ymax <- max(c(carriers_unknown_qtl$cases,
              carriers_unknown_snp$cases))

formatting_affected <- list(ylim(0, ymax),
                            theme(legend.position = "none"))

plot_affected <- ((affected_plot(carriers_unknown_snp) + 
                     formatting_affected +
                     ggtitle("Affected without balancing selection") ) /
                    (affected_plot(carriers_unknown_qtl) +
                       formatting_affected +
                       ggtitle("Affected with balancing selection")))

plot_affected
```



## The effect of exempting top carrier sires from selection

```{r}
carriers_selection_snp_10exempt <- combined_carriers(results_selection_snp_10exempt)
carriers_selection_qtl_10exempt <- combined_carriers(results_selection_qtl_10exempt)

carriers_selection_snp_100exempt <- combined_carriers(results_selection_snp_100exempt)
carriers_selection_qtl_100exempt <- combined_carriers(results_selection_qtl_100exempt)

lethal_frequency_plot(carriers_selection_snp) +
    lethal_frequency_plot(carriers_selection_qtl) +    
    lethal_frequency_plot(carriers_selection_snp_10exempt) +
    lethal_frequency_plot(carriers_selection_qtl_10exempt) +
    lethal_frequency_plot(carriers_selection_snp_100exempt) +
    lethal_frequency_plot(carriers_selection_qtl_100exempt) +
    plot_layout(ncol = 2)
```




## Figure 2: Effect of selection against lethal

```{r}

summarise_trend <- function(stats) {
    summarise(group_by(stats, generation),
              lower_g1 = quantile(mean_g1, 0.05),
              average_g1 = mean(mean_g1),
              upper_g1 = quantile(mean_g1, 0.95),
              lower_g2 = quantile(mean_g2, 0.05),
              average_g2 = mean(mean_g2),
              upper_g2 = quantile(mean_g2, 0.95))
}

genetic_trend_stats_avoid_cxc_qtl <- summarise_trend(stats_avoid_cxc_qtl)
genetic_trend_stats_avoid_cxc_snp <- summarise_trend(stats_avoid_cxc_snp)

genetic_trend_stats_selection_snp <- summarise_trend(stats_selection_snp)
genetic_trend_stats_selection_qtl <- summarise_trend(stats_selection_qtl)

genetic_trend_stats_selection_snp_10exempt <- summarise_trend(stats_selection_snp_10exempt)
genetic_trend_stats_selection_snp_100exempt <- summarise_trend(stats_selection_snp_100exempt)

genetic_trend_stats_selection_qtl_10exempt <- summarise_trend(stats_selection_qtl_10exempt)
genetic_trend_stats_selection_qtl_100exempt <- summarise_trend(stats_selection_qtl_100exempt)


genetic_trend_comparison <- rbind(transform(genetic_trend_stats_avoid_cxc_qtl,
                                            case = "no selection",
                                            selection = "balancing selection"),
                                  transform(genetic_trend_stats_selection_qtl,
                                            case = "no carrier sires",
                                            selection = "balancing selection"),
                                  transform(genetic_trend_stats_selection_qtl_10exempt,
                                            case = "allow carriers in top 10",
                                            selection = "balancing selection"),
                                  transform(genetic_trend_stats_selection_qtl_100exempt,
                                            case = "allow carriers in top 100",
                                            selection = "balancing selection"),
                                  transform(genetic_trend_stats_avoid_cxc_snp,
                                            case = "no selection",
                                            selection = "no balancing selection"),
                                  transform(genetic_trend_stats_selection_snp,
                                            case = "no carrier sires",
                                            selection = "no balancing selection"),
                                  transform(genetic_trend_stats_selection_snp_10exempt,
                                            case = "allow carriers in top 10",
                                            selection = "no balancing selection"),
                                  transform(genetic_trend_stats_selection_snp_100exempt,
                                            case = "allow carriers in top 100",
                                            selection = "no balancing selection"))


formatting_fig2 <- list(scale_colour_manual(values = c("#bdd7e7", "#6baed6", "#3182bd", "#08519c"),
                                            name = "",
                                            breaks = c("no selection",
                                                       "allow carriers in top 100",
                                                       "allow carriers in top 10",
                                                       "no carrier sires")),
                        theme_bw(),
                        theme(panel.grid = element_blank(),
                              strip.background = element_blank(),
                              legend.position = "bottom"),
                        xlab("Generation"))


## Reformat for combined trend figure

genetic_trend_comparison_g1 <- genetic_trend_comparison[, c(1:4, 8:9)]
genetic_trend_comparison_g2 <- genetic_trend_comparison[, c(1, 5:9)]

genetic_trend_comparison_g1$goal <- "Breeding goal 1"
genetic_trend_comparison_g2$goal <- "Breeding goal 2"

colnames(genetic_trend_comparison_g1)[2:4] <- c("lower", "average", "upper")
colnames(genetic_trend_comparison_g2)[2:4] <- c("lower", "average", "upper")

genetic_trend_comparison_long <- rbind(genetic_trend_comparison_g1,
                                       genetic_trend_comparison_g2)

ymax_trend <- max(genetic_trend_comparison_long$upper)
ymin_trend <- min(genetic_trend_comparison_long$lower)

fig2_trend <- ggplot(data = filter(genetic_trend_comparison_long,
                                   selection == "no balancing selection" &
                                   generation %in% c(1, seq(from = 5, to = 20, by = 5)))) +
    geom_pointrange(aes(x = generation,
                        ymin = lower,
                        y = average,
                        ymax = upper,
                        colour = case),
                    size = 0.3,
                    position = position_dodge(width = 2)) +
    ylim(ymin_trend, ymax_trend) + 
    facet_wrap(~ goal) +
    formatting_fig2 +
    theme(legend.position = "right") +
    ylab("Average genetic value") +
    ggtitle("Genetic trend")



carriers_combined_selection <- rbind(transform(carriers_avoid_cxc_snp,
                                               case = "no selection",
                                               selection = "no balancing selection"),
                                     transform(carriers_selection_snp,
                                               case = "no carrier sires",
                                               selection = "no balancing selection"),
                                     transform(carriers_selection_snp_10exempt,
                                               case = "allow carriers in top 10",
                                               selection = "no balancing selection"),
                                     transform(carriers_selection_snp_100exempt,
                                               case = "allow carriers in top 100",
                                               selection = "no balancing selection"),
                                     transform(carriers_avoid_cxc_qtl,
                                               case = "no selection",
                                               selection = "balancing selection"),
                                     transform(carriers_selection_qtl,
                                               case = "no carrier sires",
                                               selection = "balancing selection"),
                                     transform(carriers_selection_qtl_10exempt,
                                               case = "allow carriers in top 10",
                                               selection = "balancing selection"),
                                     transform(carriers_selection_qtl_100exempt,
                                               case = "allow carriers in top 100",
                                               selection = "balancing selection"))

carriers_combined_selection_summary <- summarise(group_by(carriers_combined_selection, generation, case, selection),
                                                 average = mean(carriers/n),
                                                 lower = quantile(carriers/n, 0.05),
                                                 upper = quantile(carriers/n, 0.95))

carriers_combined_selection_summary$selection <-
    factor(carriers_combined_selection_summary$selection,
           levels = c("no balancing selection", "balancing selection"))


fig2_carrier_frequency <- ggplot() +
    geom_pointrange(aes(x = generation,
                        y = average,
                        ymin = lower,
                        ymax = upper,
                        colour = case),
                    size = 0.3,
                    data = carriers_combined_selection_summary,
                    position = position_dodge(width = 1)) +
    facet_wrap(~ selection) +
    formatting_fig2 +
    ggtitle("Carrier frequency") +
    ylab("Carrier frequency")


fig2 <- 
    fig2_carrier_frequency

fig2
```


```{r}
pdf("figures/fig2_selection_against_carriers.pdf", height = 4)
print(fig2)
dev.off()
```


Carriers in generation 5:

```{r}
filter(carriers_combined_selection_summary, generation == 5)
```



Supplementary figure on the effect of balancing selection

```{r}

fig_supp_g1 <- ggplot(data = filter(genetic_trend_comparison, 
                                          generation %in% c(1, seq(from = 5, to = 20, by = 5)))) +
    geom_pointrange(aes(x = generation,
                        ymin = lower_g1,
                        y = average_g1,
                        ymax = upper_g1,
                        colour = case),
                    size = 0.3,
                    position = position_dodge(width = 2)) +
    ylim(ymin_trend, ymax_trend) + 
    facet_wrap(~ selection) +
    formatting_fig2 +
    theme(legend.position = "none") +
    ylab("Average genetic value") +
    ggtitle("Breeding goal 1")

fig_supp_g2 <- ggplot(data = filter(genetic_trend_comparison, 
                                          generation %in% c(1, seq(from = 5, to = 20, by = 5)))) +
    geom_pointrange(aes(x = generation,
                        ymin = lower_g2,
                        y = average_g2,
                        ymax = upper_g2,
                        colour = case),
                    size = 0.3,
                    position = position_dodge(width = 2)) +
    facet_wrap(~ selection) +
    ylim(ymin_trend, ymax_trend) + 
    formatting_fig2 +
    theme(legend.position = "bottom") +
    ylab("Average genetic value") +
    ggtitle("Breeding goal 2")

fig_supp_selection <- 
    (fig_supp_g1 / fig_supp_g2)

fig_supp_selection
```

```{r}
pdf("figures/suppl_fig_selection_against_carriers_balancing.pdf")
print(fig_supp_selection)
dev.off()
```



## Alternative figure for selection against carriers in the style of Fig 1

```{r}
fig2_alt_carrier_frequency <- ggplot() +
    geom_line(aes(x = generation,
                  y = carriers/n,
                  group = paste(replicate, case, selection)),
              colour = "grey",
              data = filter(carriers_combined_selection,
                            case != "no selection")) +
    geom_pointrange(aes(x = generation,
                        y = average,
                        ymin = lower,
                        ymax = upper),
                    data = filter(carriers_combined_selection_summary,
                                  case != "no selection")) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          strip.background = element_blank()) +
    facet_grid(case ~ selection) +
    ggtitle("Carrier frequency with and without balancing selection") +
    ylab("Carrier frequency") +
    xlab("Generations")

fig2_alt_carrier_frequency

```

```{r}
pdf("figures/alt_fig2_selection_against_carriers.pdf")
print(fig2_alt_carrier_frequency)
dev.off()
```




## Figure 3: Relationship between effect and frequency in balancing selection

```{r}
balancing_stats_avoid_cxc_qtl <- get_balancing_stats_divergence(results_avoid_cxc_qtl)
balancing_stats_unknown_qtl <- get_balancing_stats_divergence(results_unknown_qtl)

balancing_stats_combined <- rbind(transform(balancing_stats_avoid_cxc_qtl,
                                            case = "avoid carrier-carrier matings"),
                                  transform(balancing_stats_unknown_qtl,
                                            case = "no testing"))


balancing_stats_combined$delta_effect <- balancing_stats_combined$lethal_qtl_effect_g1 -
    balancing_stats_combined$lethal_qtl_effect_g2 
balancing_stats_combined$delta_freq <- balancing_stats_combined$top_carrier_frequency_g1 -
    balancing_stats_combined$top_carrier_frequency_g2


formatting_fig3 <- list(theme_bw(),
                        theme(panel.grid = element_blank()),
                        scale_colour_manual(values = c("#9ecae1", "#3182bd"),
                                            name = ""))

max_effect <- max(c(balancing_stats_combined$lethal_qtl_effect_g1,
                    balancing_stats_combined$lethal_qtl_effect_g2))
max_freq <- max(c(balancing_stats_combined$top_carrier_frequency_g1,
                  balancing_stats_combined$top_carrier_frequency_g2))


fig3_effect_g1 <- qplot(x = lethal_qtl_effect_g1,
                        y = top_carrier_frequency_g1,
                        colour = case,
                        data = balancing_stats_combined) +
    formatting_fig3 +
    geom_smooth(method = lm, se = FALSE, show.legend = FALSE) +
    theme(legend.position = "none") +
    ggtitle("Breeding goal 1") +
    ylab("Carrier frequency among top 10%") +
    xlab("Effect of lethal on breeding goal 1") +
    ylim(0, max_freq)

fig3_effect_g2 <- qplot(x = lethal_qtl_effect_g2,
                        y = top_carrier_frequency_g2,
                        colour = case,
                        data = balancing_stats_combined) +
  formatting_fig3 +
  geom_smooth(method = lm, se = FALSE, show.legend = FALSE) +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, linetype = 2) +
  ggtitle("Breeding goal 2") +
  ylab("Carrier frequency among top 10%") +
  xlab("Effect of lethal on breeding goal 2") +
  ylim(0, max_freq)

fig3_freq_scatter <- qplot(x = top_carrier_frequency_g1,
                           y = top_carrier_frequency_g2,
                           colour = case,
                           data = balancing_stats_combined) +
    ggtitle("Subpopulation carrier frequency") +
    xlab("Carrier frequency among top 10%, goal 1") +
    ylab("Carrier frequency among top 10%, goal 2") +
    xlim(0, max_freq) + 
    ylim(0, max_freq) + 
    geom_abline(slope = 1, intercept = 0, linetype = 2) +
    formatting_fig3
    

fig3 <- (fig3_effect_g1 +
           geom_vline(xintercept = 0, linetype = 2) +
           xlim(-max_effect, max_effect)) +
  (fig3_effect_g2 + xlim(-max_effect, max_effect)) +
  fig3_freq_scatter + 
  plot_layout(ncol = 2,
              guides = "collect") +
  plot_annotation(tag_levels = "a") & theme(legend.position = "bottom")

fig3
```



```{r}
pdf("figures/fig3_effect_frequency.pdf")
print(fig3)
dev.off()
```


```{r}
results_target_parameters <-
    read_results(system("ls simulations/divergence_simulations/target_parameters/avoid_carrier_x_carrier_qtl_results_*.Rds",
                        intern = TRUE))

carriers_target_parameters <- combined_carriers(results_target_parameters)

carriers_long_target_parameters <-
  data.frame(goal = c(rep("goal1", nrow(carriers_target_parameters)),
                      rep("goal2", nrow(carriers_target_parameters))),
             top_carrier_freq = c(carriers_target_parameters$top_goal1_frequency,
                                  carriers_target_parameters$top_goal2_frequency),
             replicate = c(carriers_target_parameters$replicate,
                           carriers_target_parameters$replicate),
             generation = c(carriers_target_parameters$generation,
                            carriers_target_parameters$generation))

top_carrier_summary <- summarise(group_by(carriers_long_target_parameters, generation, goal),
                                 average = mean(top_carrier_freq),
                                 lower = quantile(top_carrier_freq, 0.05),
                                 upper = quantile(top_carrier_freq, 0.95))

top_carrier_summary$Trait <- ifelse(top_carrier_summary$goal == "goal1",
                                    "goal 1",
                                    "goal 2")

fig3_target_parameters <- ggplot() +
  geom_pointrange(aes(x = generation,
                      y = average,
                      ymin = lower,
                      ymax = upper,
                      colour = Trait),
                  position = position_dodge(width = 0.5),
                  data = top_carrier_summary) +
  scale_colour_manual(values = c("grey", "black")) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Generation") +
  ylab("Carrier frequency among top 10%") +
  annotate("text", x = 4, y = 0.17, label = "Effect on goal 1 = 7\nEffect on goal 2 = 0")
  
fig3_target_parameters
```

```{r}
fig3_alt <- ((fig3_effect_g1 +
                geom_vline(xintercept = 0, linetype = 2) +
                xlim(-max_effect, max_effect)) +
               (fig3_effect_g2 + xlim(-max_effect, max_effect)) +
               plot_layout(ncol = 2,
                           guides = "collect") & theme(legend.position = "bottom")) /
  (fig3_target_parameters + plot_layout(guides = "keep")) + 
  plot_annotation(tag_levels = "a") 
```

```{r}
pdf("figures/fig3_effect_frequency_alt.pdf")
print(fig3_alt)
dev.off()
```



## Number of carriers used

```{r}
sires_used_summary <- summarise(group_by(carriers_combined_selection, case, selection, generation),
                                average = mean(carrier_sires_used),
                                lower = quantile(carrier_sires_used, 0.05),
                                upper = quantile(carrier_sires_used, 0.95))

ggplot() +
  geom_line(aes(x = generation,
                y = carrier_sires_used,
                group = replicate),
            colour = "grey",
            data = filter(carriers_combined_selection, case != "no carrier sires")) +
  geom_pointrange(aes(x = generation,
                      y = average,
                      ymin = lower,
                      ymax = upper),
                  data = sires_used_summary) +
  facet_wrap(case ~ selection, scale = "free_y") +
  theme_bw() +
  theme(panel.grid = element_blank()) 

summarise(group_by(filter(sires_used_summary, generation < 19), case, selection),
          per_generation = mean(average),
          lower = quantile(average, 0.05),
          upper = quantile(average, 0.95))


```



## Effect on the breeding goals in the first generations of selection

```{r}
stats_split_sel <- split(stats_selection_qtl, stats_selection_qtl$replicate)
stats_split_nosel <- split(stats_avoid_cxc_qtl, stats_avoid_cxc_qtl$replicate)

gain_generation1 <- function(stats_split) {
    map_dfr(seq_along(stats_split),
            function(i) data.frame(g1 = stats_split[[i]]$mean_g1[2] - stats_split[[i]]$mean_g1[1],
                                   g2 = stats_split[[i]]$mean_g2[2] - stats_split[[i]]$mean_g2[1]))
}

gain_generation1_sel <- gain_generation1(stats_split_sel)
gain_generation1_nosel <- gain_generation1(stats_split_nosel)
                  
```




